---
title: "Easy Cobb-Douglas Utility Function"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `microecon` package is a simple package allowing to easily solve and visualize Cobb-Douglas utility function given a budget constraint.

Install the package directly from **github** with

```{r, message=FALSE, error=FALSE, warning=FALSE}
library(devtools)
install_github("giacomovagni/microecon")
```

## Dependencies

To use `microecon` package please make sure you have installed the `tidyverse` and `ggthemes` libraires.

```{r, message=FALSE, error=FALSE, warning=FALSE}
library(microecon)
library(tidyverse)
library(ggthemes)
```

## A Simple Cobb-Douglas Utility Model

There are currently three main functions from the `microecon` package.

The first function, `cobbs_douglas_utility()`, solves a simple budget constraint optimization for two goods $x$ and $y$.

Imagine that you have a budget of \$100. The price of $x$ is \$1 and the price of $y$ is also \$1. 

You have the following utility function

$$U(x,y) = x^a y^b$$

with $a = 0.5$ and $b=0.5$

You can use the function as follow

```{r}
model1 = cobbs_douglas_utility(I = 100, px = 1, py = 1, a = 0.5, b = 0.5)
```

The function retrieves various elements such as the `optimal bundle`, the maximum utility given the budget `max_U`, and the predicted values for the two goods given the maximum utility `df`.

In this example, the optimal choice is $x=50$ and $y=50$.

```{r}
model1$optimal_bundle
model1$df %>% slice(5:10)
```

## Compare Cobb-Douglas models

You can compare two or more models with the function `cobbs_douglas_models()`. 

First save in a new vector two Cobb-Douglas models using `cobbs_douglas_utility()`. 

```{r}
model1 = cobbs_douglas_utility(I = 100, px = 1, py = 1)
```

In your second model the prices of $x,y$ have increase to \$ 2

```{r}
model2 = cobbs_douglas_utility(I = 100, px = 2, py = 2)
```

In your third model your budget has increased to 120.

```{r}
model3 = cobbs_douglas_utility(I = 120, px = 2, py = 2)
```


You can then visualize the two models with

```{r}
cb_models = cobbs_douglas_models(model1, model2, model3)
cb_models$fig
```

------------------------------------------------------------------------

Finally, you can find the optimal bundle for several goods with the function `cobb_douglas_mult()`. 

`a` is a vector of the returns, i.e. the power in the equation $U(x,y) = x1^{a1} x2^{a1} x3^{a1}$, $px$ is a vector of prices for each good and $I$ is the Budget ($I$).

```{r}
# for 3 goods
cobb_douglas_mult(a = c(0.1, 0.4, 0.5), px = c(1,1,1.2), I = 1000)
```
